06_DESeq_manipulation_groups

### 2019-04-09 analysis/06_DESeq2_manipulate_groups.Rmd
diff groups

---
title: "05_DESeq_manipulation"
output: md_document
---

@@

knitr::opts_chunk$set(fig.path = '../figures/manipulation/',cache=TRUE)

@@

m.colData$sextissue <- as.factor(paste(m.colData$sex, m.colData$tissue, sep = "_"))
```

@@

  dds <- DESeqDataSetFromMatrix(countData = countData,
                              colData = colData,
                              design = ~ treatment )
  dds <- dds[ rowSums(counts(dds)) > 2, ] ## pre-filter genes 
  dds <- DESeq(dds) # Differential expression analysis
  vsd <- vst(dds, blind=FALSE) # variance stabilized

#create list of groups
a <- levels(colData$treatment)
b <- levels(colData$treatment)

@@

dat$V1 <- factor(dat$V1, levels = 
                              c("m.inc.d3", "m.inc.d8", "m.inc.d9",
                                "m.inc.d17",  "prolong", "extend", "m.n2"))
dat$V2 <- factor(dat$V2, levels = 
                              c("m.inc.d3", "m.inc.d8", "m.inc.d9",
                                "m.inc.d17",  "prolong", "extend", "m.n2"))

@@

# create the dataframe using my function pcadataframe
pcadata <- pcadataframe(vsd, intgroup=c("treatment"), returnData=TRUE)
percentVar <- round(100 * attr(pcadata, "percentVar"))
percentVar

pca12 <- ggplot(pcadata, aes(PC1, PC2,color = treatment)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC1: ", percentVar[1],"% variance")) +
  ylab(paste0("PC2: ", percentVar[2],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25) +
  labs(subtitle = eachgroup)
print(pca12)

print(summary(aov(PC1 ~ treatment, data=pcadata)))
print(TukeyHSD(aov(PC1 ~ treatment, data=pcadata), which = "treatment"))

print(summary(aov(PC2 ~ treatment, data=pcadata))) 
print(TukeyHSD(aov(PC2 ~ treatment, data=pcadata), which = "treatment"))

pca34 <- ggplot(pcadata, aes(PC3, PC4,color = treatment)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC3: ", percentVar[3],"% variance")) +
  ylab(paste0("PC4: ", percentVar[4],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25) +
  labs(subtitle = eachgroup)
print(pca34)

# see http://bioconductor.org/packages/devel/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#heatmap-of-the-count-matrix
sampleDists <- dist(t(assay(vsd)))

sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- NULL
colnames(sampleDistMatrix) <- colData$treatment
colors <- colorRampPalette(brewer.pal(9, "Blues"))(255)
pheatmap(sampleDistMatrix,
         clustering_distance_rows=sampleDists,
         clustering_distance_cols=sampleDists,
         col=colors,
         fontsize = 6, 
         main = eachgroup)
}
```

>>>>

---
title: "06_DESeq_manipulation_groups"
output: md_document
---

@@

knitr::opts_chunk$set(fig.path = '../figures/manipulation_groups/',cache=TRUE)

@@

m.colData$sextissue <- as.factor(paste(m.colData$sex, m.colData$tissue, sep = "_"))

m.colData$newtreatment <- ifelse(grepl("m.inc.d3", m.colData$treatment), "remove.eggs",
                          ifelse(grepl("m.inc.d9", m.colData$treatment), "remove.eggs",
                          ifelse(grepl("m.inc.d17", m.colData$treatment), "remove.eggs",
                          ifelse(grepl("m.inc.d8", m.colData$treatment), "add.hatchlings",
                          ifelse(grepl("extend", m.colData$treatment), "add.hatchlings",
                          ifelse(grepl("prolong", m.colData$treatment), "keep.eggs",
                          ifelse(grepl("m.n2", m.colData$treatment), "remove.chicks",
                                   NA)))))))

m.colData$newtreatment <- factor(m.colData$newtreatment, levels = 
                              c("remove.eggs", "keep.eggs", "add.hatchlings", "remove.chicks"))

@@

  dds <- DESeqDataSetFromMatrix(countData = countData,
                              colData = colData,
                              design = ~ newtreatment )
  dds <- dds[ rowSums(counts(dds)) > 2, ] ## pre-filter genes 
  dds <- DESeq(dds) # Differential expression analysis
  vsd <- vst(dds, blind=FALSE) # variance stabilized

#create list of groups
a <- levels(colData$newtreatment)
b <- levels(colData$newtreatment)

numDEGs <- function(group1, group2){
  res <- results(dds, contrast = c("newtreatment", group1, group2), independentFiltering = T)
  sumpadj <- sum(res$padj < 0.1, na.rm = TRUE)
  return(sumpadj)
}

@@

dat$V1 <- factor(dat$V1, levels = 
                              c("remove.eggs", "keep.eggs", "add.hatchlings", "remove.chicks"))
dat$V2 <- factor(dat$V2, levels = 
                              c("remove.eggs", "keep.eggs", "add.hatchlings", "remove.chicks"))

@@

# create the dataframe using my function pcadataframe
pcadata <- pcadataframe(vsd, intgroup=c("newtreatment"), returnData=TRUE)
percentVar <- round(100 * attr(pcadata, "percentVar"))
percentVar

pca12 <- ggplot(pcadata, aes(PC1, PC2,color = newtreatment)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC1: ", percentVar[1],"% variance")) +
  ylab(paste0("PC2: ", percentVar[2],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25) +
  labs(subtitle = eachgroup)
print(pca12)

print(summary(aov(PC1 ~ newtreatment, data=pcadata)))
print(TukeyHSD(aov(PC1 ~ newtreatment, data=pcadata), which = "newtreatment"))

print(summary(aov(PC2 ~ newtreatment, data=pcadata))) 
print(TukeyHSD(aov(PC2 ~ newtreatment, data=pcadata), which = "newtreatment"))

pca34 <- ggplot(pcadata, aes(PC3, PC4,color = newtreatment)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC3: ", percentVar[3],"% variance")) +
  ylab(paste0("PC4: ", percentVar[4],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25) +
  labs(subtitle = eachgroup)
print(pca34)
}
```
